Hunting the Parent of the Orphan Stream: Identifying Stream Members \\ from Low-Resolution Spectroscopy

### 2012-12-08 2012-oss-mrs/ms.tex
Re-submitted this version to the journal

We have targeted K-giant members of the Orphan Stream in order to investigate their detailed chemistry. Because K-giants are difficult to unambiguously detect from photometry alone, low-resolution spectroscopy is required to estimate stellar parameters and determine radial velocities. The Orphan Stream has a low spatial over-density on the order of $\sim$10\%, which makes it difficult to separate stream members from halo stars. However, there is a well described distance gradient along the stream \citep{Belokurov_et-al_2007, Newberg_et-al_2010} which provides an indication on where we should focus our spectroscopic efforts.

@@

This dwarf/giant separation method was also employed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate selection was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807\,{\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807\,{\AA} appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars. Thus, we have used the 8807\,{\AA} Mg\,\textsc{I} selection throughout the rest of our analysis.

@@

We have presented a detailed analysis to isolate individual Orphan Stream K-giants from low-resolution spectroscopy using a combination of photometric, kinematic, gravity, metallicity, and proper motion information. Although each individual criteria is likely to induce some level contamination, their intersection reveals nine highly probable, self-consistent, Orphan Stream K-giants.  We deduce a median stream metallicity of $[\mbox{Fe/H}] = -1.63 \pm 0.19$ and find an intrinsically wide metallicity spread of $\sigma = 0.56$\,dex, indicative of a dSph origin. Unlike other stellar tracers, K-type giants can exist at all metallicities, hence our derived metallicity spread is likely representative of the true stream metallicity distribution function. Recall that the metallicity determination was performed after kinematic and gravity cuts, and three of our most probable members lay perfectly on a 10\,Gyr isochrone of $[\mbox{Fe/H}] = -1.63$. However, it is clear that more data is required to fully characterize the stream metallicity distribution function. Our data indicate a distance to the stream of $22.5 \pm 2.0$\,kpc at $(l, b) = (250\,^\circ,\,50\,^\circ)$, in agreement with that deduced by \citet{Grillmair_2006} and \citet{Newberg_et-al_2010}.

>>>>

We have targeted K-giant members of the Orphan Stream in order to investigate their detailed chemistry. Because K-giants are difficult to unambiguously detect from photometry alone, low-resolution spectroscopy is required to estimate stellar parameters and determine radial velocities. The Orphan Stream has an extremely low spatial over-density, which makes it difficult to separate stream members from halo stars. However, there is a well described distance gradient along the stream \citep{Belokurov_et-al_2007, Newberg_et-al_2010} which provides an indication on where we should focus our spectroscopic efforts.

@@

This dwarf/giant separation method was also employed using the total equivalent width of the Mg\,\textsc{I}\,b triplet lines. Both analyses were entirely consistent with each other: essentially the same candidate list was found using both techniques. However, given slightly poorer signal at the Mg\,\textsc{I}\,b triplet, we were forced to adopt many more upper limits than when using the 8807\,{\AA} line. Because we classify all upper limits as being ``non-dwarfs" (i.e. potential giants), we deduced a slightly larger candidate sample for the Mg\,\textsc{I}\,b analysis, which was primarily populated by upper limits. In conclusion, we found the Mg\,\textsc{I} line at 8807\,{\AA} appeared to be a more consistent dwarf discriminant given our weak S/N \--- particularly for our fainter stars. Thus, we have used the 8807\,{\AA} Mg\,\textsc{I} selection throughout the rest of our analysis.

@@

We have presented a detailed analysis to isolate individual Orphan Stream K-giants from low-resolution spectroscopy using a combination of photometric, kinematic, gravity, metallicity, and proper motion information. Although each individual criterion is likely to induce some level of contamination, their intersection reveals nine highly probable, self-consistent, Orphan Stream K-giants.  We deduce a median stream metallicity of $[\mbox{Fe/H}] = -1.63 \pm 0.19$ and find an intrinsically wide metallicity spread of $\sigma = 0.56$\,dex, indicative of a dSph origin. Unlike other stellar tracers, K-type giants can exist at all metallicities, hence our derived metallicity spread is likely representative of the true stream metallicity distribution function. Recall that the metallicity determination was performed after kinematic and gravity cuts, and three of our most probable members lay perfectly on a 10\,Gyr isochrone of $[\mbox{Fe/H}] = -1.63$. However, it is clear that more data is required to fully characterize the stream metallicity distribution function. Our data indicate a distance to the stream of $22.5 \pm 2.0$\,kpc at $(l, b) = (250\,^\circ,\,50\,^\circ)$, in agreement with that deduced by \citet{Grillmair_2006} and \citet{Newberg_et-al_2010}.
